---
title: "Dynamic Documents for Reproducible Research"
format: 
  html: 
   embed-resources: true
editor: visual
---

# initialisation

### loading libraries

```{r}
#| output: false
#| warning: false
library(tidyverse)
library(here)
library(Sleuth3)
library(ggbeeswarm)
library(janitor)
library(svglite)
library(patchwork)

#loading custom made functions
source(here("Functions", "Functions for Quarto Assignment.R"))
```

### loading data

```{r}
#| output: false
#| warning: false
RawData <- case0201
Data <- RawData
Data$Year <- as.factor(Data$Year)
write.csv(RawData,here("Data","RawData.csv"))
write.csv(Data,here("Data","Data.csv"))
```

# introduction

\[short paragraph explaining the analysis of aims in your own words\]

introduce the datasetnhello

### Dataset summary

```{r}
#| label: "Summary Statistics"
#| warning: false 
#| echo: false
print("------------All Beak Depth Data------------", quote = FALSE)
summary(Data$Depth)
print("", quote = FALSE)
print("----------Beak Depth Data for 1976----------", quote = FALSE)
summary(filter(Data, Year == 1976)$Depth)
print("", quote = FALSE)
print("----------Beak Depth Data for 1978----------", quote = FALSE)
summary(filter(Data, Year == 1978)$Depth)
```

```{r}
#| label: "Descriptive Plot"
#| warning: false 
#| output: false 
DesPlot <- BoxGraph(Data, Data$Depth, "Beak Depth (mm)",
                    Data$Year, "Year" , "Descriptive Plot", 
                    FALSE) 
SaveSVG(DesPlot,here("Figures", "Descriptive Plot.svg"),
        5, 7, 1.2)
```

![Figure 1: Descriptive plot of the raw data](Figures/Descriptive%20Plot.svg){width="529"}

# methods - statistical analysis

\[a short paragraph explaining the data, statistical model and analysis\]

Data appears to support use of linear model analysis...

## creation and testing assumptions of linear model

```{r}
#| label: "creation of linear model"
#| warning: false
#| output: false
Model <- lm(Depth ~ Year, data = Data)
```

```{r}
#| label: "Diagnostic Plots"
#| warning: false
#| output: false
svglite(here("Figures", "Diagnostic Plots.svg"), width = 10,
          height = 8,
          scaling = 1.3)
par(mfrow=c(2,2))
DiaPlots <- (plot(Model,1) | plot(Model,2)) / (plot(Model,3) | plot(Model,5))
dev.off()
```

![Figure 3: Diagnostic plots for the Linear model](Figures/Diagnostic%20Plots.svg){width="601" height="340"}

The

# results

```{r}
summary(Model)
```

### Linear Model Plot

change this to mean with error bars

```{r}
#| warning: false 
#| output: false 
LmPlot <- BoxGraph(Data, Data$Depth, "Beak Depth (mm)", Data$Year, "Year" , "Model Plot", TRUE) 
SaveSVG(LmPlot, here("Figures", "Linear Model Plot.svg"), 5, 7, 1.2)
```

![](Figures/Linear%20Model%20Plot.svg){width="579"}

# conclusion

\[a short paragraph reporting the results of the analysis\]

-   small table of some appropriate output

-   Figure which is purely representative of the data

-   figure including models and stuff

-   multi panel figure showing assessment of

# conclusion

maybe mention limitation of unpaired data

methods - statistical analysis

# results

# Appendix

## Contents of "Functions for Quarto Assignment.R"

I have pasted here the Functions that I have designed myself for full transparency of operation.

```{r}
#| eval: false

```
