---
Y---
title: "Dynamic Documents for Reproducible Research"
format: html
editor: visual
---

### loading libraries

```{r}
#| output: false
#| warning: false
library(tidyverse)
library(here)
library(Sleuth3)
library(ggbeeswarm)
```

### loading data

```{r}
RawData <- case0201
Data <- RawData
Data$Year <- as.factor(Data$Year)
write.csv(RawData,here("Data","RawData.csv"))
write.csv(Data,here("Data","Data.csv"))
```

# introduction

\[short paragraph explaining the analysis of aims in your own words\]

# methods - statistical analysis

\[a short paragraph explaining the data, statistical model and analysis\]

# results

\[a short paragraph reporting the results of the analysis\]

**include:**

-   small table of some appropriate output

-   Figure which is purely representative of the data

-   figure including models and stuff

-   multi panel figure showing assessment of

## representative table

## figures

### descriptive plot

```{r}
DesPlot <- ggplot(Data, aes(y = Depth, x = Year, colour = Year)) +
  theme_bw() +
  geom_boxplot(width = 0.7) +
  geom_beeswarm(dodge.width=1.5) +
  theme(axis.text.y   = element_text(size=14, color= "black"),
        axis.text.x   = element_text(size=14, color= "black", angle = 45, hjust=1),
        axis.title.y  = element_text(size=16),
        axis.title.x  = element_text(size=16),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.line = element_line(colour = "black", size=0.1),
        panel.border = element_rect(colour = "black", fill=NA, size=1),
        plot.title = element_text(hjust = 0.02, vjust = 0.1, size= 16),
        legend.key = element_blank(),
        legend.text = element_blank(),
        legend.background = element_blank()) +
  xlab('Year')+
  ylab('Depth (mm)')+
  ggtitle('Descriptive Plot')
DesPlot

```

# conclusion

-    maybe mention limitation of unpaired data
