---
title: "Dynamic Documents for Reproducible Research"
format: html
editor: visual
---

# initialisation

### loading libraries

```{r}
#| output: false
#| warning: false
library(tidyverse)
library(here)
library(Sleuth3)
library(ggbeeswarm)
library(janitor)
library(svglite)
library(patchwork)
source(here("Functions", "Functions for Quarto Assignment.R"))
```

### loading data

```{r}
#| output: false
#| warning: false
RawData <- case0201
Data <- RawData
Data$Year <- as.factor(Data$Year)
write.csv(RawData,here("Data","RawData.csv"))
write.csv(Data,here("Data","Data.csv"))
```

# introduction

\[short paragraph explaining the analysis of aims in your own words\]

# methods - statistical analysis

\[a short paragraph explaining the data, statistical model and analysis\]

# results

\[a short paragraph reporting the results of the analysis\]

-   small table of some appropriate output

-   Figure which is purely representative of the data

-   figure including models and stuff

-   multi panel figure showing assessment of

## representative table

## figures

### descriptive plot

```{r}
#| warning: false
DesPlot <- BoxGraph(Data, Data$Depth, "Beak Depth (mm)", Data$Year, "Year" , "Descriptive Plot", FALSE)
DesPlot

LmPlot <- BoxGraph(Data, Data$Depth, "Beak Depth (mm)", Data$Year, "Year" , "Model Plot", TRUE)
LmPlot

```

# conclusion

maybe mention limitation of unpaired data

methods - statistical analysis

# results

# conclusion
