---
title: "Dynamic Documents for Reproducible Research"
format: 
  html: 
   embed-resources: true
editor: visual
---

# Initialization

Preparation of packages, functions and Data. The contents of "Functions for Quarto Assignment.R" can be found in the Appendix.

### Packages and Functions

```{r}
#| label: "Loading packages and functions"
#| output: false
#| warning: false
library(tidyverse)
library(here)
library(Sleuth3)
library(ggbeeswarm)
library(janitor)
library(svglite)
library(patchwork)

#loading custom made functions
source(here("Functions", "Functions for Quarto Assignment.R"))
```

### Data

```{r}
#| label: "Loading data"
#| output: false
#| warning: false
RawData <- case0201
Data <- RawData
Data$Year <- as.factor(Data$Year)
write.csv(RawData,here("Data","RawData.csv"))
write.csv(Data,here("Data","Data.csv"))
```

# introduction

\[short paragraph explaining the analysis of aims in your own words\]

In 1977 there was a Severe drought on the Galapagos Island of Daphne Major, during this drought the only food available was a seed, larger and tougher than the medium ground finch (Geospiza fortis) usually consumes. There is a well established correlation between beak size and size/ toughness of seeds eaten in finches (Grant and Grant, 2003). This is likely due to a selective advantage of a larger beak being more capable of breaking down this tougher food-source.

This analysis will explore a dataset of beak sizes sampled before (1976) and after (1978) this drought. The aim is to test whether the captured data is consistent with the theory that natural selection has increased the size of beak depth as a response to this drought.

# methods - statistical analysis

\[a short paragraph explaining the data, statistical model and analysis\]

### Summary data

```{r}
#| label: "Printing of summary statistics"
#| warning: false 
#| echo: false
D76 <- filter(Data, Year == 1976)
D78 <- filter(Data, Year == 1978)
message("------------All Beak Depth Data-------------", quote = FALSE)
summary(Data$Depth)
message("----------Beak Depth Data for 1976----------", quote = FALSE)
summary(D76$Depth)
message("----------Beak Depth Data for 1978----------", quote = FALSE)
summary(D78$Depth)
```

### Descriptive plot

```{r}
#| label: "Defining and saving descriptive plots"
#| warning: false 
#| output: false 
DesPlot <- BoxGraph(Data, Data$Depth, "Beak Depth (mm)",
                    Data$Year, "Year" , "Descriptive Plot", 
                    FALSE) 

SaveSVG(DesPlot,here("Figures", "Descriptive Plot.svg"),
        5, 7, 1.2)
# svglite(here("Figures", "Histograms.svg"), width = 10,
#           height = 8,
#           scaling = 1.3)
# Hist1 <- Histogram(Data, Data$Depth, "Beak Depth (mm)", "All 
#                    Data", 1)
# Hist2 <- Histogram(D76, D76$Depth, "Beak Depth (mm)", "1976
#                    Data", 2)
# Hist3 <- Histogram(D78, D78$Depth, "Beak Depth (mm)", "All 
#                    Data", 3)
# combined_plot <- (Hist1 / (Hist2 | Hist3)) + 
#   plot_layout(heights = c(2, 1))
# dev.off()
svglite(here("Figures", "Histograms.svg"), width = 10,
          height = 8,
          scaling = 1.3)
par(mfrow =c(2,2))
Hist1 <- Histogram(Data, Data$Depth, "Beak Depth (mm)", "All 
                   Data", 1)
Hist2 <- Histogram(D76, D76$Depth, "Beak Depth (mm)", "1976
                   Data", 2)
Hist3 <- Histogram(D78, D78$Depth, "Beak Depth (mm)", "All 
                   Data", 3)
Histograms <- (Hist1 / (Hist2 | Hist3))
dev.off()
```

![Figure 1: Descriptive plot of the raw data](Figures/Descriptive%20Plot.svg){width="553"}

The Data appears to be

-   assymetrical to the right

-   linear model comparing beak depth with year

-   seeing whether it has significantly increased

-   null and alternative hypothesis

## creation and testing assumptions of linear model

```{r}
#| label: "Creation of linear model"
#| warning: false
#| output: false
Model <- lm(Depth ~ Year, data = Data)
```

```{r}
#| label: "Defining and saving of diagnostic plots"
#| warning: false
#| output: false
svglite(here("Figures", "Diagnostic Plots.svg"), width = 10,
          height = 8,
          scaling = 1.3)
par(mfrow=c(2,2))
DiaPlots <- (plot(Model,1) | plot(Model,2)) / (plot(Model,3) | plot(Model,5))
dev.off()
```

![Figure 2: Diagnostic plots for the Linear model](Figures/Diagnostic%20Plots.svg){width="676" height="527"}

The diagnostic plots suggest that none of the assumptions of the linear model have been violated and it is suitable for our analysis.

# results

### Linear Model Plot

```{r}
#| label: "Defining and saving the linear model plot"
#| warning: false  
#| output: false  
LmPlot <- BoxGraph(Data, Data$Depth, "Beak Depth (mm)", 
                   Data$Year, "Year" , "Model Plot", TRUE)
SaveSVG(LmPlot, here("Figures", "Linear Model Plot.svg"), 
        5, 7, 1.2)
```

![Figure 3: Graph showing the linear model compared to the data](Figures/Linear%20Model%20Plot.svg){width="579"}

```{r}
summary(Model)
```

# conclusion

\[a short paragraph reporting the results of the analysis\]

-   small table of some appropriate output

-   Figure which is purely representative of the data

-   figure including models and stuff

-   multi panel figure showing assessment of

# conclusion

maybe mention limitation of unpaired data

methods - statistical analysis

# Appendix

## Bibliography

Grant, R.B., Grant, P.R., 2003. What Darwin’s Finches Can Teach Us about the Evolutionary Origin and Regulation of Biodiversity. BioScience 53, 965–975. [https://doi.org/10.1641/0006-3568(2003)053\[0965:WDFCTU\]2.0.CO;2](https://doi.org/10.1641/0006-3568(2003)053[0965:WDFCTU]2.0.CO;2)

## Contents of "Functions for Quarto Assignment.R"

I have pasted here the Functions that I have designed myself for full transparency of operation.

```{r}
#| eval: false
#Boxplot creation
BoxGraph <- function(Dataset, Yaxis, Ytitle, Xaxis, Xtitle, 
                     Title, Stat){
  Summary <- Dataset %>%
    group_by(Year) %>%
    summarise(
      Mean = mean(Depth),
      Lower = mean(Depth) - qt(0.975, df = n() - 1) * sd(Depth) / sqrt(n()),
      Upper = mean(Depth) + qt(0.975, df = n() - 1) * sd(Depth) / sqrt(n())
    )
  
  plot <- ggplot(Dataset, aes(y = Yaxis, x = Xaxis, 
                              colour = Xaxis)) +
    theme_bw() +
    (if (Stat != TRUE) {
      geom_boxplot(width = 0.6)
    }) +
    geom_beeswarm(dodge.width = 1.5, alpha = 0.5) +
    (if (Stat == TRUE) {
      geom_smooth(method = "lm", aes(group = 1), 
                  se = FALSE, linetype = "solid", colour = "#606060")
    }) +
    (if (Stat == TRUE) {
      geom_errorbar(data = Summary, aes(x = Year, 
                                        ymin = Lower, ymax = Upper),
                    width = 0.08, size = 0.6, color = "#606060", inherit.aes = FALSE) 
    }) +
    (if (Stat == TRUE) {
      stat_summary(fun = "mean", geom = "point", size = 4, color = "black", shape = 18) 
    }) +
    theme(axis.text.y   = element_text(size = 14, color = "black"),
          axis.text.x   = element_text(size = 14, color = "black", 
                                       angle = 45, hjust = 1),
          axis.title.y  = element_text(size = 16),
          axis.title.x  = element_text(size = 16),
          panel.background = element_blank(),
          axis.line = element_line(colour = "black", size = 0.1),
          panel.border = element_rect(colour = "black", fill = NA,
                                      size = 1),
          plot.title = element_text(hjust = 0.02, vjust = 0.1,
                                    size = 16),
          legend.position = "none") +
    xlab(Xtitle) +
    ylab(Ytitle) +
    ggtitle(Title)
  return(plot)
}
#saving graphs as SVG for improved compatability
SaveSVG <- function(data, file, Height, Width, Scaling){
  svglite(file, width = Width,
          height = Height,
          scaling = Scaling)
  print(data)
  dev.off()
}
```
